---
title: "Implausible temperature records"
author: "Lincoln Ngo"
date: "04/01/2020"
output: 
  html_document:
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE, message=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The data below shows the **maximum temperature** recorded on the given dates at weather stations around the Sydney metropolitan area. Maximum temperatures in red are unlikely to be correct given their large deviation from other maximum temperatures recorded on the same date at other stations around Sydney. For this reason, it may be appropriate to remove these values from the historical data. All data shown is in degrees Celsius (°C).  

<br>

The respective station names and ID numbers for the weather stations included in the table below are:

* SYDNEY (OBSERVATORY HILL) (station ID: 066062)
* SYDNEY AIRPORT AMO (station ID: 066037)
* BANKSTOWN AIRPORT AWS (station ID: 066137)
* PARRAMATTA NORTH (MASONS DRIVE) (station ID: 066124)
* LIVERPOOL (WHITLAM CENTRE) (station ID: 067035)
* CANTERBURY RACECOURSE AWS (station ID: 066194)

<br>

The table below was produced using the R programming language. All code has been included in this document for transparency and so that the result can be reproduced.


```{r read_data, message = FALSE}
library(kableExtra)
library(tidyverse)
library(bomrang)


#download the data
bankstown = get_historical(stationid = "066137", type = "max") %>% as.data.frame()
sydney = get_historical(stationid = "066062", type = "max") %>% as.data.frame()
airport = get_historical(stationid = "066037", type = "max") %>% as.data.frame()
liverpool = get_historical(stationid = "067035", type = "max") %>% as.data.frame()
parramatta = get_historical(stationid = "066124", type = "max") %>% as.data.frame()
canterbury = get_historical(stationid = "066194", type = "max") %>% as.data.frame()


#get_temp() extracts the maximum temperature on the date provided
get_temp = function(df, dd, mm, yyyy) {
  temp = df %>%
    filter(day == dd & month == mm & year == yyyy) %>%
    pull(max_temperature)
  if (length(temp) == 0) {
    return(NA)
  } else {
      return(temp)
  }

}

#dates of suspicious temperature records
days = c(12, 13, 9, 24, 11, 4, 19, 4, 15, 12)
months = c(8, 11, 3, 3, 8, 11, 9, 4, 12, 12)
years = c(1971, 1987, 1993, 1993, 1993, 1994, 1999, 2005, 2006, 2008)


#new data frame that contains data from all six stations
combined_temps = data.frame(day = days,
                            month = months,
                            year = years,
                            sydney = NA,
                            syd_airport = NA,
                            bankstown = NA,
                            parramatta = NA,
                            liverpool = NA,
                            canterbury = NA                            
                            )

for (i in 1:nrow(combined_temps)) {
  
  current_d = combined_temps$day[i]
  current_m = combined_temps$month[i]
  current_y = combined_temps$year[i]
  
  combined_temps$sydney[i] = get_temp(sydney, 
                                      current_d, 
                                      current_m, 
                                      current_y)
  combined_temps$syd_airport[i] = get_temp(airport,
                                           current_d,
                                           current_m,
                                           current_y)
  combined_temps$canterbury[i] = get_temp(canterbury,
                                          current_d,
                                          current_m,
                                          current_y)
  combined_temps$bankstown[i] = get_temp(bankstown,
                                         current_d,
                                         current_m,
                                         current_y)
  combined_temps$parramatta[i] = get_temp(parramatta,
                                          current_d,
                                          current_m,
                                          current_y)
  combined_temps$liverpool[i] = get_temp(liverpool,
                                         current_d,
                                         current_m,
                                         current_y)
  
}

#row names of the combined data set are dates
rownames(combined_temps) = c("12 August 1971",
                             "13 November 1987",
                             "9 March 1993",
                             "24 March 1993",
                             "11 August 1993",
                             "4 November 1994",
                             "19 September 1999",
                             "4 April 2005",
                             "15 December 2006",
                             "12 December 2008")

#highlight implausible temperature readings and fade NA values
combined_temps[is.na(combined_temps$canterbury), "canterbury"] = cell_spec(combined_temps[is.na(combined_temps$canterbury), "canterbury"],
                                                                           "html",
                                                                           color = "LightGrey")
combined_temps[is.na(combined_temps$liverpool), "liverpool"] = cell_spec(combined_temps[is.na(combined_temps$liverpool), "liverpool"],
                                                                         "html",
                                                                         color = "LightGrey"
)

combined_temps[c(2,7), "sydney"] = cell_spec(combined_temps[c(2,7), "sydney"],
                                             "html",
                                             color = "Red")
combined_temps[-c(2,7), "bankstown"] = cell_spec(combined_temps[-c(2,7), "bankstown"],
                                                 "html",
                                                 color = "Red")

#table of data
combined_temps[,-c(1:3)] %>% 
  kable(escape = FALSE, align = c("rrrrrr")) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))
  
  

```